Accurate Binding of Sodium and Calcium to a POPC Bilayer by Effective Inclusion of Electronic Polarization

### 2018-03-13 Manuscript/SI.tex
Updates to repy reviwers' comments

@@

\section{Ternary complex model in simulations}
The NMR data about PC headgroup order parameters
and atomic absorption spectroscopy data were previously
best explained using a ternary complex binding model~\cite{altenbach84}.
In this model, one calcium is assumed to form
complexes with two lipids, i.e. with the binding stoichiometry of~1~\ce{Ca^{2+}}:2~POPC.
The model predicts a linear relationship between quantities 
$C_b$ and $\sqrt{C_b/C_I}$, where $C_b$ is the
mole fraction of bound \ce{Ca^{2+}} per POPC
and $C_I$ is the concentration of free cations at the plane of ion binding~\cite{altenbach84}.
Experimentally detemined $C_b$ %was obtained from an extrapolation of linear relation 
from NMR measurements and atomic absorption spectroscopy %for low concetrations of \ce{CaCl2}.
%Such an extrapolation is valid as long as the mode of \ce{Ca^{2+}} binding 
%remains constant throughout the extrapolation range. 
together with $C_I$ calculated from the Poisson-Boltzmann equation
gave a good agreement with the predictions of the
ternary complex model~\cite{altenbach84}.

>>>>

\newpage
\section{Electronic continuum correction applied to CHARMM36 model}
In order to demonstrate that ECC can be applied also to other than Lipid14 model,
we scaled the charges and Lennard-Jones radii of headgroup, glycerol backbone and
carbonyl atoms of the CHARMM36 model of POPC~\cite{klauda10}. We used the
scaling factors optimized for Lipid14 in the main text, i.e., 0.8 for charges and 0.89 for Lennard-Jones
radius. Changes of the headgroup order parameters as a function of added ions
and number density profiles along membrane normal from the ECC-CHARMM model are shown 
in Figs. \ref{scaledCHARMM} and \ref{scaledCHARMMdensity}, respectively,
together with the results from CHARMM36 simulations \cite{catte16,charmmPOPC450mMCaClfiles}
and experiments \cite{akutsu81,altenbach84}. As shown previously \cite{catte16},
the binding of Ca$^{2+}$ is significantly overestimated in CHARMM36 simulations.
In ECC-CHARMM simulations, the Ca$^{2+}$ binding affinity 
is significantly reduced (Fig. \ref{scaledCHARMMdensity}) and the headgroup order parameter response to the added CaCl$_2$
 is in better agreement with experiments (Fig. \ref{scaledCHARMM}).
In line with the results in main text,
Fig. \ref{scaledCHARMMcontacts} also shows faster equilibration time
for Ca$^{2+}$ binding in ECC-CHARMM model than in the original CHARMM36.

In conclusion, applying ECC to CHARMM36 simulation
improved the ion binding affinity to a lipid bilayer, even thought the
scaling factors optimized for Lipid14 were used. Also the area per molecule in
ECC-CHARMM simulation without ions (61 $\AA^2$) is only slightly
lower than the experimental value (64.3 $\AA^2$). 
The optimal scaling factors for CHARMM36 are expected to be different
than for Lipid14 because the initial partial charges are not the same
in different models. Our results indicate that the performance of ion
binding in CHARMM36, and probably in other lipid models as well, could
be further improved by finding the optimal scaling factors for the partial
charges and Lennard-Jones radii.

\newpage

\begin{figure}[!h]
  \centering
  \includegraphics[width=16.0cm]{../Fig/OP.pdf}
  \caption{\label{scaledCHARMM}
    Changes of the headgroup order parameters as a function of NaCl (left) and CaCl$_2$ (right)
    from simulations with CHARMM36 \cite{klauda10}, simulations with ECC-CHARMM parameters and experiments \cite{akutsu81,altenbach84}. 
    CHARMM36 force field parameters for simulations with and without NaCl were taken from CHARMM-GUI \cite{lee15}.
    These were ran 320 ns and 280 ns, respectively, and the first 100 ns was omitted from the analysis.
    Results for CHARMM36 with CaCl$_2$ were calculated from the last 100 ns from the data available at Ref.~\citenum{charmmPOPC450mMCaClfiles}.
    ECC-CHARMM simulations with NaCl and CaCl$_2$ were ran 260 ns and 600 ns, respectively, and the first 100 ns were omitted from the analysis.
    Sodium, calcium, and chloride ions were modeled as ECC-ions in these simulations with parameters from Refs. \citenum{martinek17, Pluharova2014}.
    All the other simulation parameters in simulations in this figure were equal to the available dataset at Ref.~\citenum{charmmPOPC450mMCaClfiles}.
    The temperature in all simulations was set to 310 K.
  }
\end{figure}

\newpage
\begin{figure}[!h]
  \centering
  \includegraphics[width=10.0cm]{../Fig/density2.pdf}
  \caption{\label{scaledCHARMMdensity}
    Number density profiles along membrane normal from CHARMM36 \cite{klauda10} and ECC-CHARMM models with NaCl and CaCl$_2$.
    The values for the relative surface excess are also shown.
  }
\end{figure}

\newpage
\begin{figure}[!h]
  \centering
  \includegraphics[width=16.0cm]{../Fig/contacts_ca.pdf}
  \caption{\label{scaledCHARMMcontacts}
    Number of bound Ca$^{2+}$ during simulations.
  }
\end{figure}

@@

\newpage
\section{Ternary complex model in simulations}
The NMR data about PC headgroup order parameters
and atomic absorption spectroscopy data were previously
best explained using a ternary complex binding model~\cite{altenbach84}.
In this model, one calcium is assumed to form
complexes with two lipids, i.e. with the binding stoichiometry of~1~\ce{Ca^{2+}}:2~POPC.
The model predicts a linear relationship between quantities 
$C_b$ and $\sqrt{C_b/C_I}$, where $C_b$ is the
mole fraction of bound \ce{Ca^{2+}} per POPC
and $C_I$ is the concentration of free cations at the plane of ion binding~\cite{altenbach84}.
Experimentally detemined $C_b$ %was obtained from an extrapolation of linear relation 
from NMR measurements and atomic absorption spectroscopy %for low concetrations of \ce{CaCl2}.
%Such an extrapolation is valid as long as the mode of \ce{Ca^{2+}} binding 
%remains constant throughout the extrapolation range. 
together with $C_I$ calculated from the Poisson-Boltzmann equation
gave a good agreement with the predictions of the
ternary complex model~\cite{altenbach84}.

### 2018-03-13 Manuscript/manuscript.tex
Updates to repy reviwers' comments

\begin{table}[tb!]
  \caption{Bulk concentrations of Ca$^{2+}$ ($C_{ion}$), relative surface excess of calcium with respect to water ($\Gamma_{Ca}^{\rm water}$),
    and the percetages of Ca$^{2+}$ bound to phosphate or carbonyl oxyges ($r^\mathrm{Ca^{2+}} _\mathrm{PO_4} $ and $r^\mathrm{Ca^{2+}} _\mathrm{O_{carb.}}$)
    in different POPC bilayer models. All systems have the same molar concentration of Ca$^{2+}$ with respect to water ($C_{ion}'$=350 mM).
  \label{tab:binding}}
  \begin{tabular}{l|c c | c | c c}
    model                  & $C_{ion}'$ & $C_{ion}\,/\,\mathrm{mM}$ & $\Gamma_{Ca}^{\rm water} \,/\, \mathrm{nm}^{-2}$  & $r^\mathrm{Ca^{2+}} _\mathrm{PO_4} $ & $r^\mathrm{Ca^{2+}} _\mathrm{O_{carb.}} $ \\
    \hline
    ECC-POPC             &  350  &  $280\pm 10 $  &  $0.06 \pm 0.01 $                           &  99\%  &    32\%    \\
    Lipid14/\AA{}qvist     &  350  &  $210\pm 10 $  &  $0.13 \pm 0.01 $                          & 100\%  &    37\%     \\
    Lipid14/Dang           &  350  &  $160\pm 10 $  &  $0.23 \pm 0.03 $                            & 100\%  &    14\%    \\
    Lipid14/ECC-ions       &  350  &  $120\pm 100$  &  $0.35 \pm 0.11 $                         & 100\%  &    23\%    \\
  \end{tabular}
\end{table}

>>>>

\begin{table}[tb!]
  \caption{Concentration of Ca$^{2+}$ with respect to water ($C_{ion}'$), bulk concentrations of Ca$^{2+}$ ($C_{ion}$), relative surface excess of calcium with respect to water ($\Gamma_{Ca}^{\rm water}$),
    and the percetages of Ca$^{2+}$ bound to phosphate or carbonyl oxyges ($r^\mathrm{Ca^{2+}} _\mathrm{PO_4} $ and $r^\mathrm{Ca^{2+}} _\mathrm{O_{carb.}}$)
    in different POPC bilayer models. 
  \label{tab:binding}}
  \begin{tabular}{l|c c | c | c c}
    model                  & $C_{ion}'$ & $C_{ion}\,/\,\mathrm{mM}$ & $\Gamma_{Ca}^{\rm water} \,/\, \mathrm{nm}^{-2}$  & $r^\mathrm{Ca^{2+}} _\mathrm{PO_4} $ & $r^\mathrm{Ca^{2+}} _\mathrm{O_{carb.}} $ \\
    \hline
    ECC-POPC             &  350  &  $280\pm 10 $  &  $0.06 \pm 0.01 $                           &  99\%  &    32\%    \\
    Lipid14/\AA{}qvist     &  350  &  $210\pm 10 $  &  $0.13 \pm 0.01 $                          & 100\%  &    37\%     \\
    Lipid14/Dang           &  350  &  $160\pm 10 $  &  $0.23 \pm 0.03 $                            & 100\%  &    14\%    \\
    Lipid14/ECC-ions       &  350  &  $120\pm 100$  &  $0.35 \pm 0.11 $                         & 100\%  &    23\%    \\
  \end{tabular}
\end{table}
